In the previous section we analyzed the design of the original RSA
model, spelled out the relation and distinction between belief, goal
and action, and proposed a familiy of alternative models based on
different interpretation of these notions. This gives rise to a series
of interesting empirical questions regarding the underlying
assumptions of the models. In order to gain insight into these
questions by comparing the predictive power of different models, we
conducted the following experiment to collect empirical data, which we
will use to test the predictions of different models in the next
section.

\subsection*{Participants}

We recruited 1032 US participants via Amazon's Mechanical Turk. Each
participant played a one-shot referential game either as a speaker or
as a listener and received a payment of 5 cents. If a participant
contributed multiple trials, only the first was included. All
participants passed the attention check described below.

\subsection*{Materials and Procedure}

Each participant saw a context consisting of three objects (Figure
\ref{exp_context} as an example) and reported on the number of objects
in the context as an attention check. Participants in the listener
conditions were told to imagine someone was talking to them who used
some utterance (depending on the specific condition tested) to refer
to exactly one of the objects in the context. Then they were asked to
choose what they thought to be the intended referent. Participants in
the speaker conditions were told to imagine they were talking to
someone and they wanted to refer to one of the objects (indicated by a
small red arrow\footnote{Participants were told that the arrow was
  only for illustration and thus the person they were talking to could
  not see it.}). Then they were asked to choose between two words to
refer to it.

\begin{figure}[htb] 
  \centering
  \begin{tikzpicture}
     \path (-1.4, 0) node [shape=rectangle, draw=green!65 ,fill= green!65 ,minimum size=28]{}
           ( 0, 0) node  [shape=circle, draw=green!65, fill= green!65,minimum size=30] {}
           ( 1.4, 0) node [shape=circle, draw=blue!65, fill=blue!65  ,minimum size=30] {}
     ;
           
  \end{tikzpicture}
  
  \caption{A sample context}\label{exp_context}
\end{figure}

Each context had a square and a circle sharing the same color, and
another circle having a different color. Hence each object had two
features: shape (square/circle) and color (green/blue) and each
context had two ambiguous (shared by two objects) features. In the
listener salience condition, participants were told that the person
talking to them was using a foreign language they could not
understand, and in the other two listener conditions, the utterances
used were the ambiguous feature of shape and color, respectively. In
the speaker conditions, the target could be any of the three objects:
the one with the unique color, the one with the unique shape and the
one with both features shared. The two words for the participants to
choose between were the features of color and shape of the target
object.

In order to minimize the effect of the confounding factors, we
counterbalanced the positions and colors of the objects, as well as
the orders of the candidate words in the speaker conditions. Thus the
unique color in a context could be either blue (as in
Fig. \ref{exp_context}) or green and the order of the objects could be
any of the permutations.\footnote{Due to the scale of our preliminary
  experiment we did not counterbalance the shape of the objects, i.e.,
  a context always had one square and two circles.}

\subsection*{Design}

There are four major design choices in our own experiment that differ
from the original experiment in \cite{Frank}. First of all, the
original experiment used a \emph{betting paradigm}, i.e., participants
were asked to bet over a range of options and they were instructed
that the amount of money that is bet on an option should reflect their
confidence that the option is correct. Even though the betting
paradigm has the merit of providing us with graded reponses from each
individual participant, the caveat is that it is unclear whether it
measures beliefs or actions. This can lead to confusion when we are to
fit the model predictions to the empirical data without knowing
whether they are directly comparable. In addition, since the betting
paradigm is more or less introspective in nature, it tends to be not
very accurate. Thus we used a \emph{forced choice design} instead,
which clearly measures the action, does not rely on introspection and,
technically speaking, provides for a straightforward likelihood
function to be used in subsequent statistical model comparison.
Secondly, since we decided to investigate the influence of the
speaker's preference as well as the listener's perceptual salience, we
focused on contexts equivalent to Figure \ref{context}, which require
reasoning that is highly reminiscent of Gricean accounts of scalar
implicature calculation, and examined different features separately.
Thirdly, we only included the features of color and shape, with binary
values (green/blue and square/circle) for two reasons. On the one hand, 
we wanted to ensure the vocabulary to be common knowledge, so we excluded 
the feature of texture. On the other hand, we excluded feature values red 
and cloud as they are significantly more salient than the others, which 
might override the pragmatic effect. 
Finally, we did not use dotted lines surrounding an object as
the way to indicate the target to the speaker, as done in \cite{Frank},
because that might unduly emphasize the feature of shape and thus be a
confound in the experiment. Instead we used a small arrow pointing to
the target.

Despite these differences, the remaining aspects in our experiment
were almost identical to those of the original one, e.g. the phrasing
of the instructions.

The results of the speaker and listener conditions of our experiment
are shown as follows.

 
\subsection{Speaker Conditions}

There were 432 participants in the speaker conditions, 144 in each condition. The numbers of participants choosing each word in each condition are shown in Table \ref{table:speaker}. Note that when the target was the object with the unique shape (as in the first row of the table), the feature of shape (``Square'') should be the optimal utterance because the listener could uniquely identify it. Similarly when the target was the object with the unique color (the second row), the optimal utterance would be the feature of color (``Blue'' in this case). When the target was the object with both features shared, both features should be equally ambiguous because of the context's symmetric nature.

\begin{table}[htb]   
  \caption{Speaker Conditions}
  \centering
  \begin{tabular}{c|ccccc}
   Target  & ``Green'' & ``Square'' & ``Blue'' & ``Circle'' & Total\\ 
     \hline
\textcolor{green!65}{\large{$\blacksquare$}}    & $9$        &   $135$   & -- & -- &144\\
\textcolor{blue!65}{\Huge{$\bullet$}}           & --        &   --      & $119$ & $25$ & 144\\
\textcolor{green!65}{\Huge{$\bullet$}}          & $63$        &   --    &  --   & $81$             &144
  \end{tabular}

  \label{table:speaker}
\end{table}

From the data in Table \ref{table:speaker}, we can see that the
speakers tend to choose the optimal feature more often when the target
has the unique shape than when it has the unique color ($\chi^2=8.5,
p<.01$). Even though they seem to prefer the feature of shape when
both of the target's features are shared, the difference is not
statistically significant from uniform random choice ($\chi^2=2.25,
p=.13$).


\subsection{Listener Conditions}

There were 600 participants in the listener conditions, 240 in the salience condition and 180 in each of the remaining two conditions. The numbers of participants choosing each object in each condition are shown in Table \ref{table:listener}. 

\begin{table}[htb] 
  \caption{Listener Conditions}
  \centering
\begin{tabular}{c|cccc}
   & \quad \textcolor{green!65}{\Large{$\blacksquare$}}&  \textcolor{green!65}{\Huge{$\bullet$}}& \textcolor{blue!65}{\Huge{$\bullet$}} & Total\\ 
     \hline
 Salience   \quad  & \quad  $71$      \quad    &   $30$    \quad     & $139$ \quad     & $240$      \\
 ``Green''   \quad  & \quad  $65$      \quad    &   $115$    \quad     & $0$ \quad     & $180$     \\
 ``Circle''   &  \quad $1$        &    $62$        & $117$ \quad     & $180$
  \end{tabular}

  \label{table:listener}
\end{table}

The result of the salience condition will be used as the empirical
estimation of contextual salience. For the other two conditions, the
object with both features shared (the green circle in the above table,
but note that we counterbalanced colors) is what the Gricean pragmatic
account predicts to be the target. The result in Table
\ref{table:listener} shows that when the message is ``Green'',
listeners prefer the green circle which is predicted by the Gricean
pragmatics, while they tend to stick to the blue circle which is more
perceptually salient when the message is ``Circle''. The behavioral
patterns in both conditions are significantly different from uniform
random choice, and they significantly differ from each other in
whether they conform to the predictions by Gricean pragmatic theory.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% TeX-PDF-mode: t
%%% End: